---
title: "Texas Campaign Data"
author: "Mark Hand"
date: "6/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=FALSE, cache = TRUE, message=FALSE, warning=FALSE)
```

```{r packages, include=FALSE}
# Cleaning up 
# rm(list=setdiff(ls(), c("all2018", "all2016")))
# rm(list=ls())

# Packages 
library(tidyverse)
library(lubridate)

```

## Where is this data from? 

This data all comes from the FEC website, which (as of 14 June 2018) was available [here](https://www.fec.gov/data/advanced/?tab=bulk-data). I used: 

* Contributions by individuals to candidates (and PACs? Double-check) (itcont)
    * 2017-2018: https://www.fec.gov/files/bulk-downloads/2018/indiv18.zip
* Contributions to candidates by other committees (e.g., PACs) (itpas2)
    * 2017-2018: https://www.fec.gov/files/bulk-downloads/2018/pas218.zip
* Any transaction from one committee to another (not involving a candidate) (itoth)
    * 2017-2018: https://www.fec.gov/files/bulk-downloads/2018/oth18.zip
* Candidate master (cn)
    * 2017-2018: https://www.fec.gov/files/bulk-downloads/2018/cn18.zip
* Committee master (cm)
    * 2017-2018: https://www.fec.gov/files/bulk-downloads/2018/cm18.zip
* Candidate-committee linkages (ccl)  
    * 2017-2018: https://www.fec.gov/files/bulk-downloads/2018/ccl18.zip

Together, I believe those represent all of of the publicly available contribution data. Becuase the data sets are so large and tab-delimited, I imported them into Excel, exported them as .csv files, imported them into the RProj and then deleted them from my computer. If you want to replicate this, either start from the RProj or download your own copy from the FEC website. 

### 2017-2018 data import
```{r packages and 2017-2018 data import}
#2017-2018: Contributions by individuals
itcont2018 <- read.table("/Users/markchand/Downloads/indiv18/itcont.txt", sep = "|", fill=TRUE, header=FALSE)
ind_names <- c("CMTE_ID","AMNDT_IND","RPT_TP","TRANSACTION_PGI","IMAGE_NUM","TRANSACTION_TP","ENTITY_TP","NAME","CITY","STATE","ZIP_CODE","EMPLOYER","OCCUPATION","TRANSACTION_DT","TRANSACTION_AMT","OTHER_ID","TRAN_ID","FILE_NUM","MEMO_CD","MEMO_TEXT","SUB_ID")
colnames(itcont2018) = ind_names
head(itcont2018)
itcont2018 %>% 
  mutate(
    IMAGE_NUM = as.numeric(IMAGE_NUM),
    ZIP_CODE = as.numeric(ZIP_CODE),
    TRANSACTION_AMT = as.numeric(TRANSACTION_AMT),
    FILE_NUM = as.numeric(FILE_NUM),
    SUB_ID = as.numeric(SUB_ID)     
         ) -> itcont2018
# Doesn't work. Figure out dates. 
# itcont2018 %>% 
#   mutate(TRANSACTION_DT = as_date(TRANSACTION_DT)) -> itcont2018

# #2017-2018: Contributions [by committees] to candidates
# itpas2018 <- read.table("/Users/markchand/Downloads/itpas2.txt", sep = "|", fill=TRUE, header=FALSE) 
# itpas2names <- c("CMTE_ID","AMNDT_IND","RPT_TP","TRANSACTION_PGI","IMAGE_NUM","TRANSACTION_TP","ENTITY_TP","NAME","CITY","STATE","ZIP_CODE","EMPLOYER","OCCUPATION","TRANSACTION_DT","TRANSACTION_AMT","OTHER_ID","CAND_ID","TRAN_ID","FILE_NUM","MEMO_CD","MEMO_TEXT","SUB_ID")
# colnames(itpas2018) = itpas2names
# head(itpas2018)

# #2017-2018: Any transaction from one committee to another
# itoth2018 <- read.table("/Users/markchand/Downloads/itoth.txt", sep = "|", fill=TRUE, header=FALSE) 
# itothnames <- c("CMTE_ID","AMNDT_IND","RPT_TP","TRANSACTION_PGI","IMAGE_NUM","TRANSACTION_TP","ENTITY_TP","NAME","CITY","STATE","ZIP_CODE","EMPLOYER","OCCUPATION","TRANSACTION_DT","TRANSACTION_AMT","OTHER_ID","TRAN_ID","FILE_NUM","MEMO_CD","MEMO_TEXT","SUB_ID")
# colnames(itoth2018) = itothnames
# head(itoth2018)

# Didn't work. Come back to this. Though for the purposes of this analysis, we're really just interested in individual donors. 
# itall2018 <- bind_rows(itcont2018, itpas2018, itoth2018, by="CMTE_ID")

# Candidates 
candidates2018 <- read.table("/Users/markchand/Downloads/cn.txt", header=FALSE, sep = "|", fill=TRUE)
can.names <- c("CAND_ID","CAND_NAME","CAND_PTY_AFFILIATION","CAND_ELECTION_YR","CAND_OFFICE_ST","CAND_OFFICE","CAND_OFFICE_DISTRICT","CAND_ICI","CAND_STATUS","CAND_PCC","CAND_ST1","CAND_ST2","CAND_CITY","CAND_ST","CAND_ZIP")
colnames(candidates2018) = can.names
candidates2018 %>% 
  mutate(
    CAND_ELECTION_YR = as.factor(CAND_ELECTION_YR)) %>% 
  rename(CMTE_ID = CAND_PCC) -> candidates2018
head(candidates2018)

# Committees 
committees2018 <- read.table("/Users/markchand/Downloads/cm.txt", header=FALSE, sep = "|", fill=TRUE)
com.names <- c("CMTE_ID","CMTE_NM","TRES_NM","CMTE_ST1","CMTE_ST2","CMTE_CITY","CMTE_ST","CMTE_ZIP","CMTE_DSGN","CMTE_TP","CMTE_PTY_AFFILIATION","CMTE_FILING_FREQ","ORG_TP","CONNECTED_ORG_NM","CAND_ID")
colnames(committees2018) = com.names
head(committees2018)
# Looks like there's something going on with CAND_ID here. 
committees2018 %>% dplyr::select(-CAND_ID) -> committees2018

# Links between candidates and committees
links2018 <- read.table("/Users/markchand/Downloads/ccl.txt", header=FALSE, sep = "|", fill=TRUE)
links.names <- c("CAND_ID","CAND_ELECTION_YR","FEC_ELECTION_YR","CMTE_ID","CMTE_TP","CMTE_DSGN","LINKAGE_ID")
colnames(links2018) = links.names
links2018 %>% 
  mutate(
   CAND_ELECTION_YR = as.factor(CAND_ELECTION_YR)) -> links2018
head(links2018)

# Joining candidates and committees 
candidates2018.2 <- left_join(candidates2018, links2018)
committees2018.2 <- left_join(committees2018, candidates2018.2)
# Handful of weird things going on here, still. Maybe a dozen entries without CMTE_ID. 
# test <- itcont2018[1:10000,]
#test.2 <- left_join(test, committees2018.2)
all2018 <- left_join(itcont2018, committees2018.2)
```

### 2015-2016 Data Import 
```{r 2015-2016 import (rework)}
#2015-2016: Contributions by individuals
itcont2016 <- read.table("/Users/markchand/Downloads/indiv16/itcont.txt", sep = "|", fill=TRUE, header=FALSE)
ind_names <- c("CMTE_ID","AMNDT_IND","RPT_TP","TRANSACTION_PGI","IMAGE_NUM","TRANSACTION_TP","ENTITY_TP","NAME","CITY","STATE","ZIP_CODE","EMPLOYER","OCCUPATION","TRANSACTION_DT","TRANSACTION_AMT","OTHER_ID","TRAN_ID","FILE_NUM","MEMO_CD","MEMO_TEXT","SUB_ID")
colnames(itcont2016) = ind_names
head(itcont2016)
itcont2016 %>% 
  mutate(
    IMAGE_NUM = as.numeric(IMAGE_NUM),
    ZIP_CODE = as.numeric(ZIP_CODE),
    TRANSACTION_AMT = as.numeric(TRANSACTION_AMT),
    FILE_NUM = as.numeric(FILE_NUM),
    SUB_ID = as.numeric(SUB_ID)     
         ) -> itcont2016
itcont2016 %>% 
  mutate(TRANSACTION_DT = as_date(TRANSACTION_DT)) -> itcont2016
# Something wrong here, but shouldn't affect the 2018 v 2016 analysis 

#2015-2016: Contributions [by committees] to candidates
# itpas2016 <- read.table("/Users/markchand/Downloads/itpas2.txt", sep = "|", fill=TRUE, header=FALSE) 
# itpas2names <- c("CMTE_ID","AMNDT_IND","RPT_TP","TRANSACTION_PGI","IMAGE_NUM","TRANSACTION_TP","ENTITY_TP","NAME","CITY","STATE","ZIP_CODE","EMPLOYER","OCCUPATION","TRANSACTION_DT","TRANSACTION_AMT","OTHER_ID","CAND_ID","TRAN_ID","FILE_NUM","MEMO_CD","MEMO_TEXT","SUB_ID")
# colnames(itpas2016) = itpas2names
# head(itpas2016)

#2015-2016: Any transaction from one committee to another
# itoth2016 <- read.table("/Users/markchand/Downloads/itoth.txt", sep = "|", fill=TRUE, header=FALSE) 
# itothnames <- c("CMTE_ID","AMNDT_IND","RPT_TP","TRANSACTION_PGI","IMAGE_NUM","TRANSACTION_TP","ENTITY_TP","NAME","CITY","STATE","ZIP_CODE","EMPLOYER","OCCUPATION","TRANSACTION_DT","TRANSACTION_AMT","OTHER_ID","TRAN_ID","FILE_NUM","MEMO_CD","MEMO_TEXT","SUB_ID")
# colnames(itoth2016) = itothnames
# head(itoth2016)

#Joining fail 
# itall2016 <- bind_rows(itcont2016, itpas2016, itoth2016)

#2015-2016 Candidates 
candidates2016 <- read.table("/Users/markchand/Downloads/cn.txt", header=FALSE, sep = "|", fill=TRUE)
can.names <- c("CAND_ID","CAND_NAME","CAND_PTY_AFFILIATION","CAND_ELECTION_YR","CAND_OFFICE_ST","CAND_OFFICE","CAND_OFFICE_DISTRICT","CAND_ICI","CAND_STATUS","CAND_PCC","CAND_ST1","CAND_ST2","CAND_CITY","CAND_ST","CAND_ZIP")
colnames(candidates2016) = can.names
candidates2016 %>% 
  mutate(
    CAND_ELECTION_YR = as.factor(CAND_ELECTION_YR)) %>% 
  rename(CMTE_ID = CAND_PCC) -> candidates2016
head(candidates2016)

#2015-2016 Committees 
committees2016 <- read.table("/Users/markchand/Downloads/cm.txt", header=FALSE, sep = "|", fill=TRUE)
com.names <- c("CMTE_ID","CMTE_NM","TRES_NM","CMTE_ST1","CMTE_ST2","CMTE_CITY","CMTE_ST","CMTE_ZIP","CMTE_DSGN","CMTE_TP","CMTE_PTY_AFFILIATION","CMTE_FILING_FREQ","ORG_TP","CONNECTED_ORG_NM","CAND_ID")
colnames(committees2016) = com.names
head(committees2016)
committees2016 %>% dplyr::select(-CAND_ID) -> committees2016

#2015-2016 Links between candidates and committees
links2016 <- read.table("/Users/markchand/Downloads/ccl.txt", header=FALSE, sep = "|", fill=TRUE)
links.names <- c("CAND_ID","CAND_ELECTION_YR","FEC_ELECTION_YR","CMTE_ID","CMTE_TP","CMTE_DSGN","LINKAGE_ID")
colnames(links2016) = links.names
links2016 %>% 
  mutate(
   CAND_ELECTION_YR = as.factor(CAND_ELECTION_YR)) -> links2016
head(links2016)

#2015-2016 Joining 
candidates2016.2 <- left_join(candidates2016, links2016)
committees2016.2 <- left_join(committees2016, candidates2016.2)
# Handful of weird things going on here, still. Maybe a dozen entries without CMTE_ID. 
# test <- itcont2016[1:10000,]
#test.2 <- left_join(test, committees2016.2)
all2016 <- left_join(itcont2016, committees2016.2)
```

### Create Texas 

### Remove non-essential dataframes 
```{r}
rm(candidates2016, candidates2016.2, candidates2018, candidates2018.2, committees2016, committees2016.2, committees2018, committees2018.2, itcont2016, itcont2018, links2016, links2018)
rm(can.names, com.names, ind_names, links.names)
```

## God Save Texas 
Let's take a look first at the folks running in Texas, and how much money they've raised from individuals: 

```{r Texas}
# Creating Texas databases 
all2018 %>% 
  filter(CAND_OFFICE_ST=="TX" | STATE =="TX") -> TX2018

all2016 %>% 
  filter(CAND_OFFICE_ST=="TX" | STATE =="TX") -> TX2016

# Removing larger dataframes 
rm(all2016, all2018)

# Checking House candidates
TX2018 %>% 
  group_by(CAND_NAME) %>% 
  summarise(TOTAL = sum (TRANSACTION_AMT)) -> TX2018.totals

# District 7 Totals
TX2018 %>% 
  filter(CAND_OFFICE_DISTRICT=="07" | CAND_OFFICE_DISTRICT=="7") %>% 
  group_by(CAND_NAME) %>% 
  summarise(TOTAL = sum (TRANSACTION_AMT)) -> TX2018.07.totals

TX2018 %>% filter(CAND_NAME, starts_with("C"))

table(TX2018$CAND_OFFICE_DISTRICT)
```






Who donated in 2017-2018 that didn't donate in 2015-2016? 
Which candidates are doing the best job of pulling in new donors? 

